---
title: "Section 7 quality - analyses for Evansen et al."
author: "Megan Evansen, Ya-Wei Li, and Jacob Malcom"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    code_folding: hide
    css: custom.css
    df_print: kable
    fig_caption: yes
    fig_width: 7
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float: true
---

# Data import

```{r steup}
library(AICcmodavg)
library(ggplot2)
library(ggthemes)
library(glmulti)
library(knitr)
library(lm.beta)
library(lmtest)
library(MASS)
library(nnet)
library(readxl)
library(readr)
library(section7.quality)

# Load the data
data(combo, package = "section7.quality")
data(formal, package = "section7.quality")
data(informal, package = "section7.quality")
data(no_prog, package = "section7.quality")
data(all_no_prog, package = "section7.quality")
```

# Analyses

We start broad and slowly drill down on the results:

1. overall quality for formal + informal
2. overall specificity of conservation actions for formals
3. components of specificity and quality for formals
4. components of informal quality

## Overall quality

### All consultations

```{r overall_qual}
all_overQ <- run_glm_combo(combo)
kable(aictab(all_overQ$mods))
```

Models 9 (best) & 2 differ only by year...base inference on 9, but note year 
effect estimates. Now to calculate the confidence intervals for the odds ratios:

```{r overall_qual_2}
mod9_OR_CI <- get_ORs_CIs(all_overQ$mods[[9]])
kable(mod9_OR_CI)
```

## Quality components: formal consultations

### Species status

```{r}
statusQ_mods <- run_polrs(formal, formal$status_Q)
aictab(statusQ_mods$mods)
statusQ_mods$summaries[[2]] 
statusQ_mods$coef[[2]]
``` 

Using model 2 b/c all params are present and very similar.

### Baseline conditions

```{r}
baselineQ_mods <- run_polrs(formal, formal$baseline_Q)
aictab(baselineQ_mods$mods)
baselineQ_mods$summaries[[1]]
baselineQ_mods$coef[[1]]
```

### effects analysis quality

```{r}
effQ_mods <- run_polrs(formal, formal$eff_Q)
aictab(effQ_mods$mods)
effQ_mods$summaries[[1]]
effQ_mods$coef[[1]]
```

### cumulative effects

```{r}
CEQ_mods <- run_polrs(formal, formal$CE_Q)
aictab(CEQ_mods$mods)
# CR_mods$summaries[[1]]
# CM_mods$coef[[1]]
```

## Quality components: informal consultations

### Action mention

```{r}
action_mods <- inf_binom(informal, informal$mention_action)
aictab(action_mods$mods)
action_mods$summaries[[1]] # nothing significant (or even close)
```

### Action analysis

```{r}
act_analysis_mods <- inf_binom(informal, informal$act_analysis)
aictab(act_analysis_mods$mods)
act_analysis_mods$summaries[[1]] # only duration significant, and tiny effect (1.07 OR)

ggplot(data = informal, aes(x = total_duration, y = act_analysis)) +
geom_jitter(height = 0.1, size = 3, alpha = 0.2) +
theme_hc()
```

### Species analysis

```{r}
sp_analysis_mods <- inf_binom(informal, informal$sp_analysis)
aictab(sp_analysis_mods$mods)
sp_analysis_mods$summaries[[2]] # only duration significant, and tiny effect (1.07 OR)

ggplot(data = informal, aes(x = total_duration, y = sp_analysis)) +
geom_jitter(height = 0.1, size = 3, alpha = 0.2) +
theme_hc()
```

### Reason informal

```{r}
mention_reason_inf_mods <- inf_binom(informal, informal$mention_reason_inf)
aictab(mention_reason_inf_mods$mods)
mention_reason_inf_mods$summaries[[1]]

qplot(data = informal, x = Service, y = mention_reason_inf, geom="violin") +
    geom_jitter(height = 0.1, width = 0.4, alpha = 0.2, size = 3) +
    theme_hc()
```

### Map included?

```{r}
map_mods <- inf_binom(informal, informal$map)
aictab(map_mods$mods)
map_mods$summaries[[1]] # Service has OR = 5.08
map_mods$summaries[[3]]

qplot(data = informal, x = Service, y = map, geom="violin") +
    geom_jitter(height = 0.1, width = 0.4, alpha = 0.2, size = 3) +
    theme_hc()
```

# Plots

## Overall quality terms

### Formal + informal

```{r}
par(mfrow = c(1, 3))
boxplot(combo$Overall_Q ~ combo$Programmatic,
        main = "Programmatics",
        ylab = "% overall quality")
boxplot(combo$Overall_Q ~ combo$Service,
        main = "With programmatics",
        ylab = "")
boxplot(combo[combo$Programmatic == "0", ]$Overall_Q ~ combo[combo$Programmatic == "0", ]$Service,
        main = "Without programmatics",
        ylab = "")
par(mfrow = c(1, 1))
```

### Formal only

```{r}
par(mfrow = c(1, 3))
boxplot(formal$Overall_Q ~ formal$Programmatic,
        main = "Programmatics",
        ylab = "% overall quality")
boxplot(formal$Overall_Q ~ formal$Service,
        main = "With programmatics",
        ylab = "")
boxplot(formal[formal$Programmatic == "0", ]$Overall_Q ~ formal[formal$Programmatic == "0", ]$Service,
        main = "Without programmatics",
        ylab = "")
par(mfrow = c(1, 1))
```

### Year effects

```{r}
par(mfrow = c(3, 1))
boxplot(combo$Overall_Q ~ factor(combo$Year),
        main = "All through time",
        ylab = "% overall quality")
boxplot(formal$Overall_Q ~ factor(formal$Year),
        main = "Formal through time",
        ylab = "% overall quality")
boxplot(informal$Overall_Q ~ factor(informal$Year),
        main = "Informals through time",
        ylab = "% overall quality")
par(mfrow = c(1, 1))
```

### Service effect

```{r}
boxplot(informal$Overall_Q ~ informal$Service,
        main = "Informals By Service",
        ylab = "% overall quality")
```

### Duration effects

```{r}
plot(combo$Overall_Q ~ combo$total_duration,
     xlab = "Duration (days)",
     ylab = "% overall quality",
     pch = c(17, 19)[as.numeric(factor(combo$Formal))],
     col = c("red", "blue")[as.numeric(factor(combo$Service))])
abline(lm(Overall_Q ~ total_duration, data = combo[combo$Service == "NMFS",]),
       col="red")
abline(lm(Overall_Q ~ total_duration, data = combo[combo$Service == "FWS",]),
       col="blue")
```

And some quick plots of quality components so I could visualize the data while
troubleshooting the models

```{r}
plt <- ggplot(data = formal, aes(status_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(baseline_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(eff_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(CE_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt
```
