---
title: "Section 7 quality - analyses for Evansen et al."
author: "Megan Evansen, Ya-Wei Li, and Jacob Malcom"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    code_folding: hide
    css: custom.css
    df_print: kable
    fig_caption: yes
    fig_width: 7
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float: true
---

# Data import

```{r steup}
library(AICcmodavg)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(glmulti)
library(knitr)
library(lm.beta)
library(lmtest)
library(MASS)
library(nnet)
library(ordinal)
library(readxl)
library(readr)
library(section7.quality)
library(viridis)

# Load the data
data(combo, package = "section7.quality")
data(formal, package = "section7.quality")
data(informal, package = "section7.quality")
data(no_prog, package = "section7.quality")
data(all_no_prog, package = "section7.quality")

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(
      seq(1, cols * ceiling(numPlots/cols)),
      ncol = cols, nrow = ceiling(numPlots/cols)
    )
  }

  if (numPlots==1) {
    print(plots[[1]])
  } else {
    grid.newpage()
    pushViewport(
      viewport(
        layout = grid.layout(
          nrow(layout), 
          ncol(layout)
        )
      )
    )

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(
        which(layout == i, arr.ind = TRUE)
      )
      print(
        plots[[i]], 
        vp = viewport(
          layout.pos.row = matchidx$row,
          layout.pos.col = matchidx$col
        )
      )
    }
  }
}
```

# Analyses

We start broad and slowly drill down on the results:

1. overall quality for formal + informal
2. overall specificity of conservation actions for formals
3. components of specificity and quality for formals
4. components of informal quality

## Overall quality

### All consultations

```{r overall_qual}
all_overQ <- run_glm_combo(combo)
kable(aictab(all_overQ$mods))

bar <- ggplot(combo, aes(combo$Overall_Q, fill = Service)) + 
  geom_histogram(
    alpha = 0.8, 
    position = position_dodge(), 
    bins = 7) + 
  labs(x = "Overall Quality") +
  scale_fill_viridis(discrete = TRUE) +
  theme_hc() +
  theme(
    legend.position = c(.05, .95),
    legend.justification = c("left", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
  )

box <- ggplot(combo, aes(y = combo$Overall_Q, x = Service)) +
  geom_boxplot(aes(fill = combo$Service), alpha = 0.8) + 
  labs(x = "Overall Quality") +
  scale_color_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(x = "", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(2.5,0.3,0.1,1.1), "cm"))

multiplot(box, bar)

# pdf(file = "inst/figures/Fig1a.pdf", height = 7, width = 7)
# multiplot(box, bar)
# dev.off()
```

Models 9 (best) & 2 differ only by year...base inference on 9, but note year 
effect estimates. Now to calculate the confidence intervals for the odds ratios:

```{r overall_qual_2}
mod9_OR_CI <- get_ORs_CIs(all_overQ$mods[[9]])
kable(mod9_OR_CI)
```

### Formal and informal consultations separate

```{r overall_qual_formal}
formal_overQ <- run_glm_combo_formal(formal)
kable(aictab(formal_overQ$mods))
mod2_OR_CI <- get_ORs_CIs(formal_overQ$mods[[2]])
kable(mod2_OR_CI)

informal_overQ <- run_glm_combo_informal(informal)
kable(aictab(informal_overQ$mods))
mod2_OR_CI <- get_ORs_CIs(informal_overQ$mods[[2]])
kable(mod2_OR_CI)

#######################
# plot!
f_bar <- ggplot(formal, aes(formal$Overall_Q, fill = Service)) + 
  geom_histogram(
    alpha = 0.8, 
    position = position_dodge(), 
    bins = 7) + 
  labs(x = "Overall Quality", y = "# consultations") +
  scale_fill_viridis(discrete = TRUE) +
  theme_hc() +
  theme(
    legend.position = c(.05, .95),
    legend.justification = c("left", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
  )

i_bar <- ggplot(informal, aes(informal$Overall_Q, fill = Service)) + 
  geom_histogram(
    alpha = 0.8, 
    position = position_dodge(), 
    bins = 7) + 
  labs(x = "Overall Quality") +
  scale_fill_viridis(discrete = TRUE) +
  theme_hc() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "none",
    legend.justification = c("left", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
  )

f_box <- ggplot(formal, aes(y = formal$Overall_Q, x = Service)) +
  geom_boxplot(aes(fill = formal$Service), alpha = 0.8) + 
  labs(x = "Overall Quality") +
  scale_color_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  ggtitle("Formal") +
  labs(x = "", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(2.5,0.2,0.1,0.1), "cm"),
        plot.title = element_text(hjust = 0.5))

i_box <- ggplot(informal, aes(y = informal$Overall_Q, x = Service)) +
  geom_boxplot(aes(fill = informal$Service), alpha = 0.8) + 
  labs(x = "Overall Quality") +
  scale_color_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  ggtitle("Informal") +
  labs(x = "", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(2.5,0.3,0.1,0.5), "cm"),
        plot.title = element_text(hjust = 0.5))

multiplot(f_box, f_bar, i_box, i_bar, cols = 2)

pdf(file = "inst/figures/Fig1b.pdf", height = 7, width = 7)
multiplot(f_box, f_bar, i_box, i_bar, cols = 2)
dev.off()
```

## Quality components: formal consultations
```{r echo = FALSE}
formal$yr <- formal$Year - 2008
fws <- filter(formal, Service == "FWS")
fws_prog <- filter(formal, Service == "FWS" & Programmatic == 1)
fws_noprog <- filter(formal, Service == "FWS" & Programmatic == 0)
```

We next dive into the individuals consultation components to identify patterns
of variation. We focus on just two predictors, Service and year in the models,
which include a random effects term for the consultation. We exclude whether the
consultation was programmatic because programmatics were only done by FWS, which 
confounds parameter estimation. Instead, we provide only a simple graphical 
analyses of programmatic effects.

### Species status

```{r}
statusQ_mods <- run_polrs(formal, formal$status_Q)
aictab(statusQ_mods$mods)
statusQ_mods$summaries[[2]] 
statusQ_mods$coef[[2]]
get_ORs_CIs_alt(statusQ_mods$mods[[2]])

sp_stat <- ggplot(formal, aes(status_Q)) +
  geom_bar(
    aes(fill = Service), 
    position = position_dodge(),
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Species Status") +
  labs(x = "", y = "# consultations") +
  theme_hc() + 
  theme(legend.position = c(0.2, 0.8),
        plot.title = element_text(hjust = 0.5))

ggplot(fws, aes(status_Q, fill = factor(Programmatic))) +
  geom_histogram(bins = 6,
                 position = "identity",
                 alpha = 0.3) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Species Status Score",
       y = "# of consultations",
       fill = "Programmatic?") +
  theme_hc() + 
  theme(legend.position = "top")

tapply(fws$status_Q, INDEX = fws$Programmatic, FUN = mean, na.rm = TRUE)
``` 

The three models all have very similar AICc values, but models 1 (Service) and 
3 (year) are subsets of model 2. Even though its AICc is ~2 units higher, we use
model 2 b/c all params are present and very similar. The main result is that 
neither Service nor year have strong signals, nor (per the plot above and the 
means between yes/no) does whether FWS consultations were programmatic.

### Baseline conditions

```{r}
baselineQ_mods <- run_polrs(formal, formal$baseline_Q)
aictab(baselineQ_mods$mods)
baselineQ_mods$summaries[[1]]
baselineQ_mods$coef[[1]]
get_ORs_CIs_alt(baselineQ_mods$mods[[1]])
tapply(formal$baseline_Q, INDEX = formal$Service, FUN = mean, na.rm = TRUE)

base_q <- ggplot(formal, aes(baseline_Q)) +
  geom_bar(
    aes(fill = Service), 
    position = position_dodge(preserve = "single"),           
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Env. Baseline") +
  labs(x = "Score", y = "# consultations") +
  theme_hc() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))

ggplot(fws, aes(baseline_Q, fill = factor(Programmatic))) +
  geom_histogram(bins = 4,
                 position = "identity",
                 alpha = 0.3) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Env. Baseline Score",
       y = "# of consultations",
       fill = "Programmatic?") +
  theme_hc() + 
  theme(legend.position = "top")

tapply(fws$baseline_Q, INDEX = fws$Programmatic, FUN = mean, na.rm = TRUE)
```

Unlike the status quality, Service stands out as the best predictor of the
quality of the environmental baseline. Similarly, the figure and means indicate
that the programmatic consultations were important to increasing the env. 
baseline score for FWS.

### effects analysis quality

```{r}
effQ_mods <- run_polrs(formal, formal$eff_Q)
aictab(effQ_mods$mods)
effQ_mods$summaries[[2]]
effQ_mods$coef[[2]]
get_ORs_CIs_alt(effQ_mods$mods[[2]])
tapply(formal$eff_Q, INDEX = formal$Service, FUN = mean, na.rm = TRUE)

eff_q <- ggplot(formal, aes(formal$eff_Q)) +
  geom_bar(
    aes(fill = formal$Service), 
    width = 0.9, 
    position = position_dodge(preserve = "single"),
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Effects Analysis") +
  labs(x = "", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))

ggplot(fws, aes(eff_Q, fill = factor(Programmatic))) +
  geom_histogram(bins = 3,
                 position = "identity",
                 alpha = 0.3) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Effects Score",
       y = "# of consultations",
       fill = "Programmatic?") +
  theme_hc() + 
  theme(legend.position = "top")

tapply(fws$eff_Q, INDEX = fws$Programmatic, FUN = mean, na.rm = TRUE)
```

As with the status of the species section, the quality of the effects analysis
did not differ between Services nor, within FWS, between (non-) programmatics.

### cumulative effects

```{r}
CEQ_mods <- run_polrs(formal, formal$CE_Q)
aictab(CEQ_mods$mods)
CEQ_mods$summaries[[1]]
CEQ_mods$coef[[1]]
get_ORs_CIs_alt(CEQ_mods$mods[[1]])
tapply(formal$CE_Q, INDEX = formal$Service, FUN = mean, na.rm = TRUE)

ce_q <- ggplot(formal, aes(CE_Q)) +
  geom_bar(
    aes(fill = Service), 
    position = position_dodge(preserve = "single"),
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Cum. Effects") +
  labs(x = "Score", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))

ggplot(fws, aes(CE_Q, fill = factor(Programmatic))) +
  geom_histogram(bins = 3,
                 position = "identity",
                 alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  labs(x = "Cum. Effects Score",
       y = "# of consultations",
       fill = "Programmatic?") +
  theme_hc() + 
  theme(legend.position = "top")

tapply(fws$CE_Q, INDEX = fws$Programmatic, FUN = mean, na.rm = TRUE)
```

As with the status of the species section, the quality of cumulative effects 
analysis did not differ between Services. There was a marginal difference - 
based on simple graphical analysis - between programmatics and non-programmatics.

```{r components_combo}
pdf(file = "inst/figures/Fig2.pdf", height = 10, width = 10)
multiplot(sp_stat, base_q, eff_q, ce_q, cols = 2)
dev.off()
```

__*Conclusion*__ 
The primary cause of the difference between FWS and NMFS overall scores was 
driven by differences in the environmental baseline analysis. NMFS's scores
tended to be higher than FWS in the other quality components, and even though
not statistically significant alone, also like contributed to the overall
difference.

## Quality components: informal consultations

### Action mention

```{r}
action_mods <- inf_binom(informal, informal$mention_action)
aictab(action_mods$mods)
action_mods$summaries[[1]] # nothing significant (or even close)

infmac <- filter(informal, !is.na(mention_action))
mact <- ggplot(infmac, aes(factor(mention_action))) +
  geom_histogram(
    aes(fill = Service), 
    position = position_dodge(preserve = "single"),
    stat = "count",
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Mention Action") +
  labs(x = "", y = "# consultations") +
  scale_x_discrete(labels = c(0, 1)) +
  theme_hc() + 
  theme(legend.position = c(0.1, 0.8),
        plot.title = element_text(hjust = 0.5))
```

There is no relationship between whether the action was mentioned and Service or
consultation duration.

### Action analysis

```{r}
act_analysis_mods <- inf_binom(informal, informal$act_analysis)
aictab(act_analysis_mods$mods)
act_analysis_mods$summaries[[1]] 
get_ORs_CIs_alt(act_analysis_mods$mods[[1]])

infan <- filter(informal, !is.na(act_analysis))
analy <- ggplot(infan, aes(factor(act_analysis))) +
  geom_histogram(
    aes(fill = Service), 
    position = position_dodge(preserve = "single"),
    stat = "count",
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Action Analysis") +
  labs(x = "Score", y = "# consultations") +
  scale_x_discrete(labels = c(0, 1)) +
  theme_hc() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))
```

Only duration significant - a very slight decline with duration - and tiny 
effect size (1.02 OR).

### Species analysis

```{r}
sp_analysis_mods <- inf_binom(informal, informal$sp_analysis)
aictab(sp_analysis_mods$mods)
sp_analysis_mods$summaries[[2]]
get_ORs_CIs_alt(sp_analysis_mods$mods[[2]])

infsp <- filter(informal, !is.na(sp_analysis))
sp_an <- ggplot(infsp, aes(factor(sp_analysis))) +
  geom_histogram(aes(fill = Service), 
                 # position = "identity",
                 position = position_dodge(prespreserve = "single"),
                 stat = "count",
                 alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Species Analysis") +
  labs(x = "", y = "") +
  scale_x_discrete(labels = c(0, 1)) +
  theme_hc() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))
```

Duration X Service is significant, but the effect size is again tiny 
(OR = 1.004).

### Reason informal

```{r}
mention_reason_inf_mods <- inf_binom(informal, informal$mention_reason_inf)
aictab(mention_reason_inf_mods$mods)
mention_reason_inf_mods$summaries[[1]]
get_ORs_CIs_alt(mention_reason_inf_mods$mods[[1]])

infre <- filter(informal, !is.na(mention_reason_inf))
reason <- ggplot(infre, aes(factor(mention_reason_inf))) +
  geom_bar(
    aes(fill = Service), 
    position = position_dodge(preserve = "single"),
    stat = "count",
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Reason Informal") +
  labs(x = "Score", y = "") +
  scale_x_discrete(labels = c(0, 1)) +
  theme_hc() + 
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))
```

Only duration significant - a very slight increase of whether the reason a 
consultation was informal with duration - and tiny effect size (1.03 OR).

### Map included?

```{r}
map_mods <- inf_binom(informal, informal$map)
aictab(map_mods$mods)
map_mods$summaries[[1]] # Service has OR = 5.08
get_ORs_CIs_alt(mention_reason_inf_mods$mods[[1]])

infmap <- filter(informal, !is.na(map))
wmap <- ggplot(infmap, aes(factor(map))) +
  geom_bar(
    aes(fill = Service), 
    bins = 2,
    stat = "count",
    position = position_dodge(prespreserve = "single"),
    alpha = 0.8) +
  scale_fill_viridis(discrete = TRUE) +
  ggtitle("Map?") +
  labs(x = "Score", y = "") +
  theme_hc() + 
  scale_x_discrete(labels = c(0, 1)) +
  theme(legend.position = "none",
        plot.title = element_text(hjust = 0.5))
```

```{r combo_informal_plot}
pdf(file = "inst/figures/Fig3.pdf", height = 8, width = 12)
multiplot(mact, analy, sp_an, reason, wmap, cols = 3)
dev.off()
```

# Additional Plots

This plots section is basically obsolete because figures have been added above.

## Overall quality terms

### Formal + informal

```{r}
par(mfrow = c(1, 3))
boxplot(combo$Overall_Q ~ combo$Programmatic,
        main = "Programmatics",
        ylab = "% overall quality")
boxplot(combo$Overall_Q ~ combo$Service,
        main = "With programmatics",
        ylab = "")
boxplot(combo[combo$Programmatic == "0", ]$Overall_Q ~ combo[combo$Programmatic == "0", ]$Service,
        main = "Without programmatics",
        ylab = "")
par(mfrow = c(1, 1))

par(mfrow = c(2, 1))
ggplot(combo, aes(x = Service, y = Overall_Q)) +
  geom_boxplot() +
  coord_flip() +
  theme_hc()

```

### Formal only

```{r}
par(mfrow = c(1, 3))
boxplot(formal$Overall_Q ~ formal$Programmatic,
        main = "Programmatics",
        ylab = "% overall quality")
boxplot(formal$Overall_Q ~ formal$Service,
        main = "With programmatics",
        ylab = "")
boxplot(formal[formal$Programmatic == "0", ]$Overall_Q ~ formal[formal$Programmatic == "0", ]$Service,
        main = "Without programmatics",
        ylab = "")
par(mfrow = c(1, 1))
```

### Year effects

```{r}
par(mfrow = c(3, 1))
boxplot(combo$Overall_Q ~ factor(combo$Year),
        main = "All through time",
        ylab = "% overall quality")
boxplot(formal$Overall_Q ~ factor(formal$Year),
        main = "Formal through time",
        ylab = "% overall quality")
boxplot(informal$Overall_Q ~ factor(informal$Year),
        main = "Informals through time",
        ylab = "% overall quality")
par(mfrow = c(1, 1))
```

### Service effect

```{r}
boxplot(informal$Overall_Q ~ informal$Service,
        main = "Informals By Service",
        ylab = "% overall quality")
```

### Duration effects

```{r}
plot(combo$Overall_Q ~ combo$total_duration,
     xlab = "Duration (days)",
     ylab = "% overall quality",
     pch = c(17, 19)[as.numeric(factor(combo$Formal))],
     col = c("red", "blue")[as.numeric(factor(combo$Service))])
abline(lm(Overall_Q ~ total_duration, data = combo[combo$Service == "NMFS",]),
       col="red")
abline(lm(Overall_Q ~ total_duration, data = combo[combo$Service == "FWS",]),
       col="blue")
```

And some quick plots of quality components so I could visualize the data while
troubleshooting the models

```{r}
plt <- ggplot(data = formal, aes(status_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(baseline_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(eff_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(CE_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt
```
