---
title: "Section 7 quality - analyses for Evansen et al."
author: "Megan Evansen, Ya-Wei Li, and Jacob Malcom"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    code_folding: hide
    css: custom.css
    df_print: kable
    fig_caption: yes
    fig_width: 7
    highlight: tango
    toc: true
    toc_depth: 3
    toc_float: true
---

# Data import

```{r steup}
library(AICcmodavg)
library(dplyr)
library(ggplot2)
library(ggthemes)
library(glmulti)
library(knitr)
library(lm.beta)
library(lmtest)
library(MASS)
library(nnet)
library(ordinal)
library(readxl)
library(readr)
library(section7.quality)
library(viridis)

# Load the data
data(combo, package = "section7.quality")
data(formal, package = "section7.quality")
data(informal, package = "section7.quality")
data(no_prog, package = "section7.quality")
data(all_no_prog, package = "section7.quality")

multiplot <- function(..., plotlist=NULL, file, cols=1, layout=NULL) {
  library(grid)
  # Make a list from the ... arguments and plotlist
  plots <- c(list(...), plotlist)

  numPlots = length(plots)
  if (is.null(layout)) {
    layout <- matrix(
      seq(1, cols * ceiling(numPlots/cols)),
      ncol = cols, nrow = ceiling(numPlots/cols)
    )
  }

  if (numPlots==1) {
    print(plots[[1]])
  } else {
    grid.newpage()
    pushViewport(
      viewport(
        layout = grid.layout(
          nrow(layout), 
          ncol(layout)
        )
      )
    )

    # Make each plot, in the correct location
    for (i in 1:numPlots) {
      matchidx <- as.data.frame(
        which(layout == i, arr.ind = TRUE)
      )
      print(
        plots[[i]], 
        vp = viewport(
          layout.pos.row = matchidx$row,
          layout.pos.col = matchidx$col
        )
      )
    }
  }
}
```

# Analyses

We start broad and slowly drill down on the results:

1. overall quality for formal + informal
2. overall specificity of conservation actions for formals
3. components of specificity and quality for formals
4. components of informal quality

## Overall quality

### All consultations

```{r overall_qual}
all_overQ <- run_glm_combo(combo)
kable(aictab(all_overQ$mods))

bar <- ggplot(combo, aes(combo$Overall_Q, fill = Service)) + 
  geom_histogram(alpha = 0.8, position = position_dodge(), bins = 7) + 
  labs(x = "Overall Quality") +
  scale_fill_viridis(discrete = TRUE) +
  theme_hc() +
  theme(
    legend.position = c(.05, .95),
    legend.justification = c("left", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
  )

box <- ggplot(combo, aes(y = combo$Overall_Q, x = Service)) +
  geom_boxplot(aes(fill = combo$Service), alpha = 0.8) + 
  labs(x = "Overall Quality") +
  scale_color_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(x = "", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(2.5,0.3,0.1,1.1), "cm"))

multiplot(box, bar)
```

Models 9 (best) & 2 differ only by year...base inference on 9, but note year 
effect estimates. Now to calculate the confidence intervals for the odds ratios:

```{r overall_qual_2}
mod9_OR_CI <- get_ORs_CIs(all_overQ$mods[[9]])
kable(mod9_OR_CI)
```

### Formal and informal consultations separate

```{r overall_qual_formal}
formal_overQ <- run_glm_combo_formal(formal)
kable(aictab(formal_overQ$mods))
mod2_OR_CI <- get_ORs_CIs(formal_overQ$mods[[2]])
kable(mod2_OR_CI)

informal_overQ <- run_glm_combo_informal(informal)
kable(aictab(informal_overQ$mods))
mod2_OR_CI <- get_ORs_CIs(informal_overQ$mods[[2]])
kable(mod2_OR_CI)

#######################
# plot!
f_bar <- ggplot(formal, aes(formal$Overall_Q, fill = Service)) + 
  geom_histogram(alpha = 0.8, position = position_dodge(), bins = 7) + 
  labs(x = "Overall Quality (formal)") +
  scale_fill_viridis(discrete = TRUE) +
  theme_hc() +
  theme(
    legend.position = c(.05, .95),
    legend.justification = c("left", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
  )

i_bar <- ggplot(informal, aes(informal$Overall_Q, fill = Service)) + 
  geom_histogram(alpha = 0.8, position = position_dodge(), bins = 7) + 
  labs(x = "Overall Quality (informal)") +
  scale_fill_viridis(discrete = TRUE) +
  theme_hc() +
  theme(
    axis.title.y = element_blank(),
    legend.position = "none",
    legend.justification = c("left", "top"),
    legend.box.just = "right",
    legend.margin = margin(6, 6, 6, 6)
  )

f_box <- ggplot(formal, aes(y = formal$Overall_Q, x = Service)) +
  geom_boxplot(aes(fill = formal$Service), alpha = 0.8) + 
  labs(x = "Overall Quality") +
  scale_color_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(x = "", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        axis.text.x = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(2.5,0.3,0.1,0.1), "cm"))

i_box <- ggplot(informal, aes(y = informal$Overall_Q, x = Service)) +
  geom_boxplot(aes(fill = informal$Service), alpha = 0.8) + 
  labs(x = "Overall Quality") +
  scale_color_viridis(discrete = TRUE) +
  scale_fill_viridis(discrete = TRUE) +
  coord_flip() +
  labs(x = "", y = "") +
  theme_hc() + 
  theme(legend.position = "none",
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        plot.margin = unit(c(2.5,0.3,0.1,0.5), "cm"))

multiplot(f_box, f_bar, i_box, i_bar, cols = 2)
```

## Quality components: formal consultations
```{r echo = FALSE}
formal$yr <- formal$Year - 2008
fws_prog <- filter(formal, Service == "FWS" & Programmatic == 1)
fws_noprog <- filter(formal, Service == "FWS" & Programmatic == 0)
```

We next dive into the individuals consultation components to identify patterns
of variation. We focus on just two predictors, Service and year in the models,
which include a random effects term for the consultation. We exclude whether the
consultation was programmatic because programmatics were only done by FWS, which 
results in badly inflated odds ratios. Instead, we provide only a simple 
graphical analyses of programmatic effects.

### Species status

```{r}
statusQ_mods <- run_polrs(formal, formal$status_Q)
aictab(statusQ_mods$mods)
statusQ_mods$summaries[[1]] 
statusQ_mods$coef[[1]]
get_ORs_CIs_alt(statusQ_mods$mods[[1]])

ggplot(formal, aes(status_Q)) +
  geom_bar(data = fws_prog, fill = "red", alpha = 0.3) +
  geom_bar(data = fws_noprog, fill = "blue", alpha = 0.3) +
  theme_hc()
``` 

Using model 2 b/c all params are present and very similar.

### Baseline conditions

```{r}
baselineQ_mods <- run_polrs(formal, formal$baseline_Q)
aictab(baselineQ_mods$mods)
baselineQ_mods$summaries[[1]]
baselineQ_mods$coef[[1]]
get_ORs_CIs_alt(baselineQ_mods$mods[[1]])

ggplot(formal, aes(baseline_Q, fill = Programmatic)) +
  geom_bar(data = fws_prog, fill = "red", alpha = 0.3) +
  geom_bar(data = fws_noprog, fill = "blue", alpha = 0.3) +
  theme_hc()
```

### effects analysis quality

```{r}
effQ_mods <- run_polrs(formal, formal$eff_Q)
aictab(effQ_mods$mods)
effQ_mods$summaries[[1]]
effQ_mods$coef[[1]]
get_ORs_CIs_alt(effQ_mods$mods[[1]])
```

### cumulative effects

```{r}
CEQ_mods <- run_polrs(formal, formal$CE_Q)
aictab(CEQ_mods$mods)
CEQ_mods$summaries[[1]]
CEQ_mods$coef[[1]]
get_ORs_CIs_alt(CEQ_mods$mods[[1]])
```

## Quality components: informal consultations

### Action mention

```{r}
action_mods <- inf_binom(informal, informal$mention_action)
aictab(action_mods$mods)
action_mods$summaries[[1]] # nothing significant (or even close)
```

### Action analysis

```{r}
act_analysis_mods <- inf_binom(informal, informal$act_analysis)
aictab(act_analysis_mods$mods)
act_analysis_mods$summaries[[1]] # only duration significant, and tiny effect (1.07 OR)

ggplot(data = informal, aes(x = total_duration, y = act_analysis)) +
geom_jitter(height = 0.1, size = 3, alpha = 0.2) +
theme_hc()
```

### Species analysis

```{r}
sp_analysis_mods <- inf_binom(informal, informal$sp_analysis)
aictab(sp_analysis_mods$mods)
sp_analysis_mods$summaries[[2]] # only duration significant, and tiny effect (1.07 OR)

ggplot(data = informal, aes(x = total_duration, y = sp_analysis)) +
geom_jitter(height = 0.1, size = 3, alpha = 0.2) +
theme_hc()
```

### Reason informal

```{r}
mention_reason_inf_mods <- inf_binom(informal, informal$mention_reason_inf)
aictab(mention_reason_inf_mods$mods)
mention_reason_inf_mods$summaries[[1]]

qplot(data = informal, x = Service, y = mention_reason_inf, geom="violin") +
    geom_jitter(height = 0.1, width = 0.4, alpha = 0.2, size = 3) +
    theme_hc()
```

### Map included?

```{r}
map_mods <- inf_binom(informal, informal$map)
aictab(map_mods$mods)
map_mods$summaries[[1]] # Service has OR = 5.08
map_mods$summaries[[3]]

qplot(data = informal, x = Service, y = map, geom="violin") +
    geom_jitter(height = 0.1, width = 0.4, alpha = 0.2, size = 3) +
    theme_hc()
```

# Plots

## Overall quality terms

### Formal + informal

```{r}
par(mfrow = c(1, 3))
boxplot(combo$Overall_Q ~ combo$Programmatic,
        main = "Programmatics",
        ylab = "% overall quality")
boxplot(combo$Overall_Q ~ combo$Service,
        main = "With programmatics",
        ylab = "")
boxplot(combo[combo$Programmatic == "0", ]$Overall_Q ~ combo[combo$Programmatic == "0", ]$Service,
        main = "Without programmatics",
        ylab = "")
par(mfrow = c(1, 1))

par(mfrow = c(2, 1))
ggplot(combo, aes(x = Service, y = Overall_Q)) +
  geom_boxplot() +
  coord_flip() +
  theme_hc()

```

### Formal only

```{r}
par(mfrow = c(1, 3))
boxplot(formal$Overall_Q ~ formal$Programmatic,
        main = "Programmatics",
        ylab = "% overall quality")
boxplot(formal$Overall_Q ~ formal$Service,
        main = "With programmatics",
        ylab = "")
boxplot(formal[formal$Programmatic == "0", ]$Overall_Q ~ formal[formal$Programmatic == "0", ]$Service,
        main = "Without programmatics",
        ylab = "")
par(mfrow = c(1, 1))
```

### Year effects

```{r}
par(mfrow = c(3, 1))
boxplot(combo$Overall_Q ~ factor(combo$Year),
        main = "All through time",
        ylab = "% overall quality")
boxplot(formal$Overall_Q ~ factor(formal$Year),
        main = "Formal through time",
        ylab = "% overall quality")
boxplot(informal$Overall_Q ~ factor(informal$Year),
        main = "Informals through time",
        ylab = "% overall quality")
par(mfrow = c(1, 1))
```

### Service effect

```{r}
boxplot(informal$Overall_Q ~ informal$Service,
        main = "Informals By Service",
        ylab = "% overall quality")
```

### Duration effects

```{r}
plot(combo$Overall_Q ~ combo$total_duration,
     xlab = "Duration (days)",
     ylab = "% overall quality",
     pch = c(17, 19)[as.numeric(factor(combo$Formal))],
     col = c("red", "blue")[as.numeric(factor(combo$Service))])
abline(lm(Overall_Q ~ total_duration, data = combo[combo$Service == "NMFS",]),
       col="red")
abline(lm(Overall_Q ~ total_duration, data = combo[combo$Service == "FWS",]),
       col="blue")
```

And some quick plots of quality components so I could visualize the data while
troubleshooting the models

```{r}
plt <- ggplot(data = formal, aes(status_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(baseline_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(eff_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt

plt <- ggplot(data = formal, aes(CE_Q)) +
       geom_bar() +
       facet_grid(. ~ Service + Programmatic) +
       theme_hc()
plt
```
